---
title: "Nathan's Hotdog Eating Contest Graphic"
author: ''
date: '2022-03-31'
slug: nathans_hotdog_eating_contest_graphic
categories: []
tags: tags
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---


This was from lab 1   

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = FALSE, comment = NA, warnings = FALSE, errors = FALSE, messages = FALSE, tidy = FALSE, cache = TRUE)
```




*loaded these packages*

```{r load-packages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(extrafont)
library(here)
library(ggthemes)
library(readr)
```

## Load the main dataset  

```{r load data, echo=TRUE, message=FALSE, warning=FALSE}
library(readr)
hot_dogs <- read_csv("hot_dogs.csv", col_types = cols(
  gender = col_factor(levels = NULL))) %>% 
 #using = NULL the code orders it automatically
 mutate(post_ifoce = year >= 1997) %>%  # boolean T/F
  filter(year >= 1981 & gender == 'male')
```

### This will be used at the end, when we isolate the female contestants to stack on the barplot  

```{r load data w fem, echo=TRUE, message=FALSE, warning=FALSE}
hdm_females <- read_csv("hdm_affil.csv", 
    col_types = cols(
      gender = col_factor(levels = NULL), affiliated = col_factor(levels = NULL)
    )) %>%
  mutate(post_ifoce = year >= 1997) %>% 
  filter(year >= 1981 & gender == "female")
```

### We will be using hdm_affil primarily in these beginning steps  

```{r echo=TRUE, message=FALSE, warning=FALSE}
hdm_affil <- hot_dogs %>%
  filter(year >= 1981 & gender == "male") # 1981 greater than or equal to and pulling only male
```


```{r glimpse, echo=TRUE, message=FALSE, warning=FALSE}
glimpse(hdm_affil)

```

#### Building the barplot    


This first code chunk is from the first dataset, 'hot_dogs' which only has 'males.'  

```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(hot_dogs, aes(x = year, y = num_eaten)) + 
  geom_col(aes(fill = post_ifoce), 
           colour = "white")+ 
  labs(x = "Year", 
       y = "Hot Dogs and Buns Consumed") +
  ggtitle("Nathan's Hot Dog Eating Contest Results, 1981-2017") +
  scale_fill_discrete(name = "",
  labels=c("Pre-IFOCE", "Post-IFOCE"))

```

#### saving the plot as an item so it can be built on throughout the process  

```{r echo=TRUE, message=FALSE, warning=FALSE}
affil_plot <- hdm_affil %>% ggplot(aes(
  x = year,
  y = num_eaten)) + 
  geom_col(aes(fill = post_ifoce)) +
  labs(x = "Year",
       y = "Hot Dogs and Buns Consumed",
  title = "Nathan's Hot Dog Eating Contest Results, 1981-2017",
  subtitle = "(Male contestants only)") +
  scale_fill_manual(values = c('#E9602B','#CCB683'), name = "IFOCE-affiliation") + 
  coord_cartesian(
 xlim = c(1980, 2018), 
ylim = c(0, 80)) 
```

#### how to add a custom theme  

```{r echo=TRUE, message=FALSE, warning=FALSE}
plot_theme <- theme(plot.title = element_text(hjust = 0.5),
                     axis.text = element_text(size = 12),
                     panel.background = element_blank(),
                    
axis.line.x = 
          element_line(color = "gray80", size = 0.5),
                    
axis.ticks = 
          element_line(color = "gray80", size = 0.5),
                     
text = 
          element_text(family = "Lato")) #extrafont

affil_plot + plot_theme
```

### All combined code, preparing to add annotations  

```{r echo=TRUE, message=FALSE, warning=FALSE}
# custom tick marks  
years_to_label <- seq(from = 1981, to = 2017, by = 4)   

# add column to use for x-axis  
hd_years <- hdm_affil %>%
  distinct(year) %>%
  mutate(year_lab = ifelse(year %in% years_to_label, year, ""))

# making a fresh plot in order to detail the formatting  
# this is essentially the same as affil_plot  

nathan_plot <- ggplot(hdm_affil, aes(
  x = year, y = num_eaten)) + 
  geom_col(aes(fill = post_ifoce)) + 
  labs(x = "Year", 
       y = "Hot Dogs and Buns Consumed", 
  title = "Nathan's Hot Dog Eating Contest Results, 1981-2017", 
  subtitle = "(Male contestants only)") +
  scale_fill_manual(
    values = c('#E9602B','#CCB683'),
                    name = "IFOCE-affiliation") + plot_theme 

# save plot with edited axis scales, breaks and tick labels
nathan_plot <- nathan_plot +
  scale_y_continuous(expand = c(0, 0),
                     breaks = seq(0, 70, 10)) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = hd_years$year,
                     labels = hd_years$year_lab) + 
  coord_cartesian(xlim = c(1980, 2017), ylim = c(0, 80)) # adjust x and y axis 
```


#### The best part... Adding annotations and removing the legend  

Annotations make reading data visualizations easier and faster to read and understand   

```{r echo=TRUE, message=FALSE, warning=FALSE}
nathan_ann <- nathan_plot +
  guides(fill = FALSE) +

  coord_cartesian(xlim = c(1980, 2022), ylim = c(0, 90)) +
  # far left segment
  annotate('segment', x=1980.75, xend=2000.25, y= 30, yend=30, size=0.5, color="#CCB683")+
  
  annotate('segment', x=1980.75, xend=1980.75, y= 30, yend=28, size=0.5, color="#CCB683") +
 
   annotate('segment', x=2000.25, xend=2000.25, y= 30, yend=28, size=0.5, color="#CCB683") +
 
   annotate('segment', x=1990, xend=1990, y= 33, yend=30, size=0.5, color="#CCB683") +
 
   annotate('text', x=1990, y=36, 
        label="No MLE/IFOCE Affiliation", color="#CCB683", family="Lato", hjust=0.5, size = 3) +
# middle segment
  annotate('segment', x=2000.75, xend=2006.25, y= 58, yend=58, size=0.5, color="#2277A0") +
 
   annotate('segment', x=2000.75, xend=2000.75, y= 58, yend=56, size=0.5, color="#2277A0") +
  
  annotate('segment', x=2006.25, xend=2006.25, y= 58, yend=56, size=0.5, color="#2277A0") +
 
   annotate('segment', x=2003.5, xend=2003.5, y= 61, yend=58, size=0.5, color="#2277A0") +

    annotate('text', x=2003.5, y=65, 
        label="MLE/IFOCE\nFormer Member", color="#2277A0", family="Lato", hjust=0.5, size = 3) +
# far right segment
 
   annotate('segment', x=2006.75, xend=2017.25, y= 79, yend=79, size=0.5, color="#E9602B") +
 
   annotate('segment', x=2006.75, xend=2006.75, y= 79, yend=77, size=0.5, color="#E9602B") +
  
  annotate('segment', x=2017.25, xend=2017.25, y= 79, yend=77, size=0.5, color="#E9602B") +
  
  annotate('segment', x=2012, xend=2012, y= 82, yend=79, size=0.5, color="#E9602B") +
  
  annotate('text', x=2012, y=86, 
        label="MLE/IFOCE Current Member", color="#E9602B", family="Lato", hjust=0.5, size = 3)
```
#### The near-finished barplot; since the female contestants will be the last bit  

```{r echo=FALSE, message=FALSE, warning=FALSE}
nathan_ann
```


### lastly, adding the female contestants   

Use the data.frame created prior labeled, 'hdm_females'. THis isolates the female contestants so they can be stacked within the barplot without overlapping    

As a review, here is how the data.frame was created   

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
hdm_females <- hot_dogs %>% 
  mutate(post_ifoce = year >= 1997) %>% 
  filter(year >= 1981 & gender == "female") 
```


```{r echo=TRUE, message=FALSE, warning=FALSE}
nathan_w_females <- nathan_ann +
  # add in the female data, and manually set a fill color
  geom_col(data = hdm_females, 
           width = 0.75, 
           fill = "#F68A39") + # orange
  labs(subtitle = NULL) # no longer need the subtitle warning about male-only data, NULL removes the subtitle
nathan_w_females
```


Please ignore my off balanced annotations, for time's sake, lets just admire this reproducible code for a later time.   



