---
title: "Animal Sounds Analysis"
author: ''
date: '2022-06-07'
slug: animal_sounds_analysis
categories: []
tags: tags
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This was lab 3


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(wordbankr)
library(here)
library(RColorBrewer)
library(wesanderson)
library(ggthemes)
library(ggsci)
library(cowplot)
library(viridis)
library(forcats)
library(colorspace)
library(ggrepel)
library(ggplot2)
library(knitr)
```


```{r load data, message=FALSE, warning=FALSE}
library(readr)
sounds <- read_csv("animal_sounds_summary.csv")
```

For the sounds dataset, I'm going to throw in some code here for reference on how to do some exploratory data analysis.  

```{r}
sounds %>% 
  group_by(sound) %>%
  summarize(total_produce = sum(kids_produce)) %>% 
  knitr::kable()

```


This shows the average proportion that was recitable (?) by all the kids in the study.  

```{r}
sounds %>% group_by(sound) %>% 
  summarise(total_prop = mean(prop_produce))
```


To show how many ages are counted and what the range of ages are (in months).  

```{r echo=TRUE, message=FALSE, warning=FALSE}
sounds %>% distinct(age) %>% count() # 11
sounds %>% distinct(age) %>% range() # 8 18
sounds %>% distinct(age) %>% pull() # shows all values
```

How many kids per age?

```{r echo=TRUE, message=FALSE, warning=FALSE}
#summarize(new_column_name = function(column_name))
library(janitor)
# janitor::adorn_totals() will add a total row at the bottom
library(knitr)
sounds %>% 
  group_by(age) %>%
  summarize(total_age_months = sum(age)) %>% 
  adorn_totals() %>%
  knitr::kable()
```

This was the code I used to produce a scatterplot that shows at what ages the children in this study were able to make specific animal noises - faceted by the animal sounds.   


```{r echo=TRUE}
scatterplot_sounds <- ggplot(sounds, aes(x = age, 
  y = prop_produce)) + 
  geom_point() +
  labs(
    x = "Age (months)", 
    y = "Proportion of Children Producing") +
  facet_wrap(~sound)
scatterplot_sounds
```


This is the code for a line graph that is color coded by animal sounds, not faceted.  


```{r echo=TRUE, message=FALSE, warning=FALSE}
line_plot_sounds <- ggplot(sounds, aes(
  x = age, 
  y = prop_produce)) + 
  geom_line(aes(group = sound)) + 
  # geom_line is grouped by sound, meaning the lines will connect to each sound group point (3 different line plots vs 1 fully connected line plot)
  geom_point(aes(color = sound), size = 2) + 
  # geom_point also grouped by color by sound
  labs(x = "Age (months)", 
       y = "Proportion of Children Producing") 

line_plot_sounds
```


